---
title: "Project"
author: "Zach Morby"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output: 
  html_document:
    df_print: paged
    code_folding: "hide"
    toc: yes
    fig_caption: yes
    theme: cerulean
    toc_float: no
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())
graphics.off()
```

```{r load packages, message = FALSE, include = FALSE}
require(fpp3)
require(tidyquant)
require(stargazer)
require(kableExtra)
require(fma)
require(reshape2)
```

```{r getdata}
VarList <- c("PCEPI", "UNRATE", "EXPINF1YR", "IPMAN", "HOUST")
fred_data <- tq_get(VarList, get="economic.data", from="1982-01-01") %>% 
  mutate(Month = yearmonth(date), value = price) %>%
  select(-c(date, price)) %>% 
  as_tsibble(index = Month, key = symbol) %>%
  pivot_wider(names_from = symbol, values_from = value)
```

```{r mutate}
clean_data <- fred_data %>% select(c(PCEPI, UNRATE, EXPINF1YR, IPMAN, HOUST)) %>%
  mutate(infl = 1200*log(PCEPI/lag(PCEPI))) %>% 
  mutate(dinfl = infl - lag(infl,1)) %>% 
  mutate(dinfl12 = 100*log(PCEPI/lag(PCEPI,12)) - lag(infl,12)) %>% 
  mutate(unrate = UNRATE - lag(UNRATE)) %>% 
  mutate(expinf1yr = EXPINF1YR - lag(EXPINF1YR)) %>% 
  mutate(ipman = 1200*log(IPMAN/lag(IPMAN))) %>% 
  mutate(houst = 100*log(HOUST/lag(HOUST))) %>% 
  select(-c(PCEPI, UNRATE, EXPINF1YR, IPMAN, HOUST)) %>% 
  drop_na()
train_data <- clean_data %>% filter_index(~ "2019-12")
test_data <- clean_data %>% filter_index("2020-01" ~ .)
```

# Introduction

We are attempting find a model that can predict/forecast future inflation, using the 	Personal Consumption Expenditures: Chain-type Price Index as the marker for inflation. To predict different models, we will use unemployment rate, 1-Year Expected Inflation, Industrial Production: Manufacturing (NAICS), and New Privately-Owned Housing Units Started: Total Units data from the FRED database. 

# Variable Explanation

We chose to use the variables as potential indicators of inflation for various reasons. Unemployment rate and inflation are both indicators of how the economy is performing, so it is possible that unemployment may be a good predictor of inflation. The 1-Year Expected Inflation is a predictor that the Federal Reserve puts out that predicts the inflation for the following year. It could be possible to predict further into the future if this variable has been historically accurate. Industrial Production has also been used as an indicator for inflation. High levels of production/manufacturing can lead to higher levels of consumption, which theoretically could increase inflation. As for New Housing Started, inflation tends to have a direct effect on  the new housing market. Higher prices and higher inflation is going to lead to higher costs of building houses, and overall less houses built. Because inflation today does affect the housing starts tomorrow, it could be possible to build a model using the data for housing starts lagged by a year to predict inflation, essentially reversing the effect to predict inflation.

# Models

We used Personal Consumption Expenditures: Chain-type Price Index to create variables for inflation, differenced inflation, and differenced inflation on a monthly basis. 

Using the differenced inflation variable, we built four different models, one using each of the four afformentioned variables as potential predictors of inflation. As shown by the graph below, this particular model is a very inaccurate model. The residuals = bad. All of the other models return similar results. They are bad.   


```{r melt, include = FALSE}
data_melt <- melt(clean_data, "Month")
ggplot(data_melt, aes(Month, value)) + 
  geom_line() + 
  facet_wrap(~variable, scales = "free", ncol = 2)
```

```{r model, include = FALSE}
fit_model <- train_data %>% 
  model(
    mUNRATE = TSLM(dinfl12 ~ 1 +
                 lag(dinfl,12) + lag(dinfl,13) + lag(dinfl,14) +
                 lag(dinfl,15) + lag(dinfl,16) + lag(dinfl,17) +
                 lag(dinfl,18) + lag(dinfl,19) + lag(dinfl,20) +
                 lag(dinfl,21) + lag(dinfl,22) + lag(dinfl,23) +
                 lag(unrate,12) + lag(unrate,13) + lag(unrate,14) +
                 lag(unrate,15) + lag(unrate,16) + lag(unrate,17) +
                 lag(unrate,18) + lag(unrate,19) + lag(unrate,20) +
                 lag(unrate,21) + lag(unrate,22) + lag(unrate,23)),
    
    mEXPINF1YR = TSLM(dinfl12 ~ 1 +
                 lag(dinfl,12) + lag(dinfl,13) + lag(dinfl,14) +
                 lag(dinfl,15) + lag(dinfl,16) + lag(dinfl,17) +
                 lag(dinfl,18) + lag(dinfl,19) + lag(dinfl,20) +
                 lag(dinfl,21) + lag(dinfl,22) + lag(dinfl,23) +
                 lag(expinf1yr,12) + lag(expinf1yr,13) + lag(expinf1yr,14) +
                 lag(expinf1yr,15) + lag(expinf1yr,16) + lag(expinf1yr,17) +
                 lag(expinf1yr,18) + lag(expinf1yr,19) + lag(expinf1yr,20) +
                 lag(expinf1yr,21) + lag(expinf1yr,22) + lag(expinf1yr,23)),
    
    mIPMAN = TSLM(dinfl12 ~ 1 +
                 lag(dinfl,12) + lag(dinfl,13) + lag(dinfl,14) +
                 lag(dinfl,15) + lag(dinfl,16) + lag(dinfl,17) +
                 lag(dinfl,18) + lag(dinfl,19) + lag(dinfl,20) +
                 lag(dinfl,21) + lag(dinfl,22) + lag(dinfl,23) +
                 lag(ipman,12) + lag(ipman,13) + lag(ipman,14) +
                 lag(ipman,15) + lag(ipman,16) + lag(ipman,17) +
                 lag(ipman,18) + lag(ipman,19) + lag(ipman,20) +
                 lag(ipman,21) + lag(ipman,22) + lag(ipman,23)),
    
    mHOUST = TSLM(dinfl12 ~ 1 +
                 lag(dinfl,12) + lag(dinfl,13) + lag(dinfl,14) +
                 lag(dinfl,15) + lag(dinfl,16) + lag(dinfl,17) +
                 lag(dinfl,18) + lag(dinfl,19) + lag(dinfl,20) +
                 lag(dinfl,21) + lag(dinfl,22) + lag(dinfl,23) +
                 lag(houst,12) + lag(houst,13) + lag(houst,14) +
                 lag(houst,15) + lag(houst,16) + lag(houst,17) +
                 lag(houst,18) + lag(houst,19) + lag(houst,20) +
                 lag(houst,21) + lag(houst,22) + lag(houst,23))
  )

tidy(fit_model)
accuracy(fit_model)
```

```{r check_fit, warning = FALSE}
fit_model %>% select(mUNRATE) %>% gg_tsresiduals()
```

```{r check, include = FALSE, warning = FALSE}
fit_model %>% select(mEXPINF1YR) %>% gg_tsresiduals()
fit_model %>% select(mIPMAN) %>% gg_tsresiduals()
fit_model %>% select(mHOUST) %>% gg_tsresiduals()
```

# Adding a Model

```{r ensemble}
fit_combo <- fit_model %>% mutate(combo = (mUNRATE + mEXPINF1YR + mIPMAN + mHOUST)/4)
```

We then created a fourth ensemble model which is an average of the four other models. This should return an overall more accurate model because it can average out some of the outliers in the models.  

# Forecast

The following shows the forecast of the five models. The black line is the actual inflation rate. Each colored line represents the prediction of each different model. The shaded area is the 95% confidence interval.


```{r forecast}
fc_inflation <- fit_combo %>% forecast(new_data = test_data)
fc_inflation %>% autoplot(filter(clean_data, year(Month) > 2016), level = c(95))
```

# Accuracy

Finally, we evaluated the accuracy of each of our five models. None of the models were very accurate. They all did a poor job predicting the actual inflation in every metric we looked at. None of the models stood out as doing a better job predicting inflation than the othe models.  

```{r accuracy, include = FALSE}
accuracy(fit_combo)
accuracy(fc_inflation, clean_data)
```

